---
title: "Literature review of corporate default prediction "
author: "Benjamin Christoffersen"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  function(...) rmarkdown::html_document(..., css=c("css/jquery.dataTables.min.css", "css/style.css")):
    toc: false
---

```{r setup, include=FALSE}
options(scipen = 100, digits = 3)
```

## My rating
My rating is my subject score based on the following criteria:

* Relevance if you are building bankruptcy prediction models
* Large data samples (not just listed firms)
* Non-financial firms
* Different models, method etc. 
* Different countries or data sets that are commonly used in literature
* Credit scoring is not considered unless I find the paper relevant

Details of the rating are explaind in "My review". An NA rating means that I have not yet read the paper/document but it is on my TODO list

## Review table

```{r, echo=FALSE,message=FALSE, warning=TRUE, error=TRUE, results='asis'}
# Read in data
review_data <- read.csv("data/review_data.csv", header = T, sep = ",", quote = '"', 
                        stringsAsFactors = F, fileEncoding = "UTF-8-BOM")
actual_headers <- read.csv("data/review_data.csv", header = F, sep = ",", quote = '"', 
                           stringsAsFactors = F, nrows = 1, fileEncoding = "UTF-8-BOM")

colnames(review_data) <- actual_headers

# Replace the model abbrevations 
model_abb <- read.table("data/model_def", sep = ",", header = F, stringsAsFactors = F)
review_data$Models <- sapply(review_data$Models, function(x){
    ms <- unlist(strsplit(x, ", "))
    ms_match <- match(ms, model_abb[, 1])
    if(any(is.na(ms_match))){
        stop("Could not find ", paste0(
            "'", ms[is.na(ms_match)], "'", collapse = " "))
    }
    paste0(model_abb[ms_match, 2], collapse = "\n")
})

# Find further article information
if(F){ # Set to FALSE if data should be loaded from output folder
  source("R/google_scholar_info.R")
  google_dat <- apply(
    review_data[, c("Article name", "Publisher")], 1, function(x)
      get_google_scholar_info(x[1], x[2]))
  google_dat <- data.frame(matrix(unlist(google_dat), nrow = length(google_dat), 
                                  byrow = T, dimnames = 
                                    list(NULL, names(google_dat[[1]]))), 
                           stringsAsFactors = F)
  
  if(all(is.na(google_dat)))
    stop("Likely got 503 errors as google banned the ip for a few days")
  
  write.table(cbind(review_data[, c("Article name", "Publisher")],
                    google_dat), 
              "output/google_data", row.names = F, col.names = T)
} else {
  google_dat <- read.table("output/google_data", row.names = NULL, 
                           stringsAsFactors = F, header = T)
  
  google_dat <- 
    google_dat[match(review_data$`Article name`, google_dat$Article.name), 
               c("journal_url", "year", "n_citations", "abstract")]
}

google_dat[, c("year", "n_citations")] <- 
  sapply(google_dat[, c("year", "n_citations")], as.numeric)

# Put urls into a <a> tag
google_dat$journal_url <- ifelse(
  is.na(google_dat$journal_url), 
  rep(NA, nrow(google_dat)), 
  paste0("<a href = '", google_dat$journal_url, "' >link</a>"))

# format column names
google_dat <- plyr::rename(google_dat, 
                           c("journal_url" = "link", "n_citations" = "Citations"))

review_data <- cbind(review_data, google_dat)

# Order rows by my train and number of citations
review_data <- 
  if(!all(is.na(review_data$Citations)))
    review_data[order(review_data$`My rating (1: best, 5: worst)`, 
                      -review_data$Citations), ] else
                        review_data[order(review_data$`My rating (1: best, 5: worst)`), ]

# Order columns and define alignment
align_n_c_order <- c(
  "Article name" = "left", 
  "My rating (1: best, 5: worst)" = "right", 
  "Citations" = "right", 
  "abstract" = "left",
  "link" = "left",
  "My review" = "left", 
  "Outcome" = "left", 
  "Sample size (failure / non-failure)" = "left",
  "Models" = "left", 
  "Data source/set" = "left", 
  "Sampling" = "left",
  "year" = "right", 
  "Publisher" = "left")

split_cells <- c(
  "Article name" = 25, 
  "My rating (1: best, 5: worst)" = 20, 
  "Citations" = 10, 
  "abstract" = 50,
  "My review" = 50, 
  "Outcome" = 50, 
  "Sample size (failure / non-failure)" = 20,
  "Models" = 25, 
  "Data source/set" = 50, 
  "Sampling" = 50,
  "year" = 4, 
  "Publisher" = 20,
  "link" = 50)

review_data <- 
  review_data[, names(align_n_c_order)]
row.names(review_data) <- NULL

# Final formating of header 
review_data <- plyr::rename(review_data, c(
  "My rating (1: best, 5: worst)" = "My rating<br>(1: best, 5: worst)",
  "Sample size (failure / non-failure)" = "Sample size<br>(failure / non-failure)"
))

# Replace line breaks with br tags
is_char <- sapply(review_data, is.character)
review_data[, is_char] <- 
  sapply(review_data[, is_char], gsub, pattern = "\\s*\\n\\s*",
         replacement = "\\ \\<br\\> \\")

# We also correct for invalid dash entries
review_data[, is_char] <- 
  sapply(review_data[, is_char], gsub, pattern = "–",
         replacement = "-")

# Print table
pander::panderOptions('knitr.auto.asis', FALSE)
pander::pandoc.table(review_data,
                     keep.line.breaks = T, 
                     justify = align_n_c_order, 
                     split.table = Inf,
                     split.cells = split_cells[names(align_n_c_order)],
                     style = 'multiline')
```

<script src="js/jquery.dataTables.min.js"></script>

<script type="text/javascript">
// Sort function for NA from https://datatables.net/forums/discussion/7446/numeric-sort-with-na-in-some-cells
$.fn.dataTableExt.oSort['numeric_ignore_nan-asc'] = function(x,y) {
if (isNaN(x) && isNaN(y)) return ((x < y) ? 1 : ((x > y) ? -1 : 0));

if (isNaN(x)) return 1;
if (isNaN(y)) return -1;

x = parseFloat( x );
y = parseFloat( y );
return ((x < y) ? -1 : ((x > y) ? 1 : 0));
};

$.fn.dataTableExt.oSort['numeric_ignore_nan-desc'] = function(x,y) {
if (isNaN(x) && isNaN(y)) return ((x < y) ? 1 : ((x > y) ? -1 : 0));

if (isNaN(x)) return 1;
if (isNaN(y)) return -1;

x = parseFloat( x );
y = parseFloat( y );
return ((x < y) ? 1 : ((x > y) ? -1 : 0));
};

$(document).ready(function(){
  $("#review-table > table > colgroup").remove();
  $("#review-table > table").removeAttr('width');
  $("#review-table > table").DataTable({
    "scrollY": "1200px",
		"scrollX":        true,
		"autoWidth": false,
		"order": [[1, "asc" ], [2, "desc" ]],
    "columnDefs": [
      { "width": "18em", "targets": 0 },  // article name
      { "width": "9em", "type": "numeric_ignore_nan", "targets": 1 },   // rating
      { "width": "5em", "type": "numeric_ignore_nan", "targets": 2 },   // citations
      { "width": "40em", "targets": 3 },  // abstact
      { "width": "6em", "targets": 4 },   // link
      { "width": "40em", "targets": 5 },  // review
      { "width": "20em", "targets": 6 },  // outcome
      { "width": "20em", "targets": 7 },  // sample size
      { "width": "20em", "targets": 8 },  // models
      { "width": "40em", "targets": 9 },  // data set
      { "width": "20em", "targets": 10 }, // sampling
      { "width": "5em", "targets": 11 },  // year
      { "width": "10em", "targets": 12 }  // publisher
    ]
  });
});
</script>